Capstone Movielens Report

### 2025-05-03 reports/capstone-movielens-report.draft4.Rmd
Resumed development of the RMD report.

---
title: "Capstone Movielens Report"
author: "Azamat Kurbanaev"
date: "`r Sys.Date()`"
bibliography: references.bib
#csl: bit-numerical-mathematics.csl 
output:
  pdf_document: 
    df_print: paged
    toc: true
    fig_caption: true
    keep_tex: true
    citation_package: biblatex
colorlinks: true
linkcolor: red
citecolor: blue
header-includes:
    - \usepackage{hyperref}
    - \usepackage[natbib=true, 
                  style=numeric, 
                  backref=true, 
                  sorting=none]{biblatex}
    - \hypersetup{backref,
          pdfpagemode=Normal,
          colorlinks=true,
          implicit=false}
vignette: >
  %\VignetteIndexEntry{RMarkdown Citations - Numeric Style}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
---

```{r setup, include=TRUE, echo = FALSE}
if(!require(tidyverse))
  install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret))
  install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table))
  install.packages("data.table", repos = "http://cran.us.r-project.org")

if(!require(gtools)) 
  install.packages("gtools")
if(!require(pak)) 
  install.packages("pak")
if(!require("pacman")) 
  install.packages("pacman")

# Loading the required libraries
library(dslabs)
library(tidyverse)

library(caret)
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(Metrics)
library(recosystem)
library(scales)
library(stringr)
library(tibble)
library(tidyr)

library(rafalib)
library(gtools)
library(pak)
library(pacman)

p_load(latex2exp)

knitr::opts_chunk$set(echo = TRUE)
```

>>>>

---
title: "Capstone Movielens Report"
author: "Azamat Kurbanaev"
date: "`r Sys.Date()`"
bibliography: references.bib
#csl: bit-numerical-mathematics.csl 
envir:
  parent.frame()
output:
  pdf_document: 
    df_print: paged
    toc: true
    fig_caption: true
    keep_tex: true
    citation_package: biblatex
colorlinks: true
linkcolor: red
citecolor: blue
header-includes:
    - \usepackage{hyperref}
    - \usepackage[natbib=true, 
                  style=numeric, 
                  backref=true, 
                  sorting=none]{biblatex}
    - \hypersetup{backref,
          pdfpagemode=Normal,
          colorlinks=true,
          implicit=false}
vignette: >
  %\VignetteIndexEntry{RMarkdown Citations - Numeric Style}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
---

```{r setup, include=FALSE, echo = FALSE}
r.path <- "../r"
src.folder <- "src"
support_scripts.folder <- "support-scripts"
support_functions.folder <- "support-functions"

r.src.path <- file.path(r.path, src.folder)
support_scripts.path <- file.path(r.src.path, support_scripts.folder)
support_functions.path <- file.path(r.src.path, support_functions.folder)
## Setup -----------------------------------------------------------------------
setup_script.file_path <- file.path(support_scripts.path,
                                               "setup.R")

# main_script.file_path <- file.path(r.src.path,
#                                    "capstone-movielens.main.R")
# stop(main_script.file_path)
stopifnot(file.exists(setup_script.file_path))

source(setup_script.file_path, 
       local = knitr::knit_global())
# or sys.source("your-script.R", envir = knitr::knit_global())

## External Common Helper functions -------------------------------------
# common_helper_functions.file_path <- file.path(support_functions.path,
#                                             "common-helper.functions.R")
# source(common_helper_functions.file_path, 
#        local = knitr::knit_global())

# knitr::knit_global()
#knitr::opts_chunk$set(echo = FALSE)
```

## Plot dependency of RMSEs vs Overal Mean Rating Deviation
\
```{r}

data.frame(delta = deviation, 
           delta.RMSE = deviation.RMSE) |> 
tuning.plot(title = TeX(r'[RMSE as a function of deviation ($\delta$) from the Overall Mean Rating ($\hat{mu}$)]'),
              xname = "delta", 
              yname = "delta.RMSE", 
              xlabel = TeX(r'[$\delta$]'), 
              ylabel = "RMSE")

### 2025-05-03 reports/capstone-movielens-report.draft4.tex
Resumed development of the RMD report.

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

@@

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\begin{verbatim}
## Loading required package: tidyverse
\end{verbatim}

\begin{verbatim}
## Warning: package 'lubridate' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## -- Attaching core tidyverse packages ------------------------ tidyverse 2.0.0 --
## v dplyr     1.1.4     v readr     2.1.5
## v forcats   1.0.0     v stringr   1.5.1
## v ggplot2   3.5.1     v tibble    3.2.1
## v lubridate 1.9.4     v tidyr     1.3.1
## v purrr     1.0.2
\end{verbatim}

\begin{verbatim}
## -- Conflicts ------------------------------------------ tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
## i Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
## Loading required package: caret
\end{verbatim}

\begin{verbatim}
## Warning: package 'caret' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## Loading required package: lattice
## 
## Attaching package: 'caret'
## 
## The following object is masked from 'package:purrr':
## 
##     lift
## 
## Loading required package: data.table
\end{verbatim}

\begin{verbatim}
## Warning: package 'data.table' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'data.table'
## 
## The following objects are masked from 'package:lubridate':
## 
##     hour, isoweek, mday, minute, month, quarter, second, wday, week,
##     yday, year
## 
## The following objects are masked from 'package:dplyr':
## 
##     between, first, last
## 
## The following object is masked from 'package:purrr':
## 
##     transpose
## 
## Loading required package: gridExtra
## 
## Attaching package: 'gridExtra'
## 
## The following object is masked from 'package:dplyr':
## 
##     combine
## 
## Loading required package: logr
\end{verbatim}

\begin{verbatim}
## Warning: package 'logr' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## Loading required package: common
\end{verbatim}

\begin{verbatim}
## Warning: package 'common' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## Loading required package: gtools
\end{verbatim}

\begin{verbatim}
## Warning: package 'gtools' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## Loading required package: pak
\end{verbatim}

\begin{verbatim}
## Warning: package 'pak' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## Loading required package: pacman
\end{verbatim}

\begin{verbatim}
## Warning: package 'pacman' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## Loading required package: recommenderlab
\end{verbatim}

\begin{verbatim}
## Warning: package 'recommenderlab' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## Loading required package: Matrix
\end{verbatim}

\begin{verbatim}
## Warning: package 'Matrix' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'Matrix'
## 
## The following objects are masked from 'package:tidyr':
## 
##     expand, pack, unpack
## 
## Loading required package: arules
\end{verbatim}

\begin{verbatim}
## Warning: package 'arules' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'arules'
## 
## The following object is masked from 'package:dplyr':
## 
##     recode
## 
## The following objects are masked from 'package:base':
## 
##     abbreviate, write
## 
## Loading required package: proxy
## 
## Attaching package: 'proxy'
## 
## The following object is masked from 'package:Matrix':
## 
##     as.matrix
## 
## The following objects are masked from 'package:stats':
## 
##     as.dist, dist
## 
## The following object is masked from 'package:base':
## 
##     as.matrix
## 
## Registered S3 methods overwritten by 'registry':
##   method               from 
##   print.registry_field proxy
##   print.registry_entry proxy
## 
## Attaching package: 'recommenderlab'
## 
## The following objects are masked from 'package:caret':
## 
##     MAE, RMSE
\end{verbatim}

\begin{verbatim}
## Warning: package 'dslabs' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'cowplot' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'cowplot'
## 
## The following object is masked from 'package:lubridate':
## 
##     stamp
\end{verbatim}

\begin{verbatim}
## Warning: package 'ggthemes' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'ggthemes'
## 
## The following object is masked from 'package:cowplot':
## 
##     theme_map
\end{verbatim}

\begin{verbatim}
## Warning: package 'Metrics' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'Metrics'
## 
## The following objects are masked from 'package:caret':
## 
##     precision, recall
\end{verbatim}

\begin{verbatim}
## Warning: package 'recosystem' was built under R version 4.4.2
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'scales'
## 
## The following object is masked from 'package:purrr':
## 
##     discard
## 
## The following object is masked from 'package:readr':
## 
##     col_factor
## 
## [conflicted] Will prefer base::as.matrix over any other package.
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{require}\NormalTok{(edx.capstone.movielens.data)) pak}\SpecialCharTok{::}\FunctionTok{pak}\NormalTok{(}\StringTok{"AzKurban{-}edX{-}DS/edx.capstone.movielens.data"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: edx.capstone.movielens.data
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(edx.capstone.movielens.data)}
\NormalTok{edx }\OtherTok{\textless{}{-}}\NormalTok{ edx.capstone.movielens.data}\SpecialCharTok{::}\NormalTok{edx}
\NormalTok{final\_holdout\_test }\OtherTok{\textless{}{-}}\NormalTok{ edx.capstone.movielens.data}\SpecialCharTok{::}\NormalTok{final\_holdout\_test}

@@

\begin{verbatim}
##      userId         movieId          rating        timestamp        
##  Min.   :    1   Min.   :    1   Min.   :0.500   Min.   :7.897e+08  
##  1st Qu.:18124   1st Qu.:  648   1st Qu.:3.000   1st Qu.:9.468e+08  
##  Median :35738   Median : 1834   Median :4.000   Median :1.035e+09  
##  Mean   :35870   Mean   : 4122   Mean   :3.512   Mean   :1.033e+09  
##  3rd Qu.:53607   3rd Qu.: 3626   3rd Qu.:4.000   3rd Qu.:1.127e+09  
##  Max.   :71567   Max.   :65133   Max.   :5.000   Max.   :1.231e+09  
##     title              genres         
##  Length:9000055     Length:9000055    
##  Class :character   Class :character  
##  Mode  :character   Mode  :character  
##                                       
##                                       
## 
\end{verbatim}

@@

\begin{verbatim}
##      userId         movieId          rating        timestamp        
##  Min.   :    1   Min.   :    1   Min.   :0.500   Min.   :7.897e+08  
##  1st Qu.:18096   1st Qu.:  648   1st Qu.:3.000   1st Qu.:9.467e+08  
##  Median :35768   Median : 1827   Median :4.000   Median :1.035e+09  
##  Mean   :35870   Mean   : 4108   Mean   :3.512   Mean   :1.033e+09  
##  3rd Qu.:53621   3rd Qu.: 3624   3rd Qu.:4.000   3rd Qu.:1.127e+09  
##  Max.   :71567   Max.   :65133   Max.   :5.000   Max.   :1.231e+09  
##     title              genres         
##  Length:999999      Length:999999     
##  Class :character   Class :character  
##  Mode  :character   Mode  :character  
##                                       
##                                       
## 
\end{verbatim}

@@

\begin{verbatim}
## `summarise()` has grouped output by 'movieId'. You can override using the
## `.groups` argument.
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

@@

\subparagraph{}\label{section-1}

@@

\begin{verbatim}
## # A tibble: 5 x 5
##   userId `Pulp Fiction (1994)` `Jurassic Park (1993)` Silence of the Lambs (19~1
##    <int>                 <dbl>                  <dbl>                      <dbl>
## 1     13                     4                     NA                         NA
## 2     16                    NA                      3                         NA
## 3     17                    NA                     NA                          5
## 4     18                     5                      3                          5
## 5     19                    NA                      1                         NA
## # i abbreviated name: 1: `Silence of the Lambs (1991)`
## # i 1 more variable: `Forrest Gump (1994)` <dbl>
\end{verbatim}

@@

\begin{verbatim}
##      userId         movieId          rating        timestamp        
##  Min.   :    8   Min.   :    1   Min.   :0.500   Min.   :8.248e+08  
##  1st Qu.:18161   1st Qu.: 1022   1st Qu.:3.000   1st Qu.:9.655e+08  
##  Median :35666   Median : 2109   Median :3.500   Median :1.058e+09  
##  Mean   :35902   Mean   : 4616   Mean   :3.473   Mean   :1.051e+09  
##  3rd Qu.:53645   3rd Qu.: 3993   3rd Qu.:4.000   3rd Qu.:1.135e+09  
##  Max.   :71565   Max.   :65133   Max.   :5.000   Max.   :1.231e+09  
##     title              genres         
##  Length:1397040     Length:1397040    
##  Class :character   Class :character  
##  Mode  :character   Mode  :character  
##                                       
##                                       
## 
\end{verbatim}

@@

\begin{verbatim}
##      userId           movieId            rating        timestamp        
##  59269  :   5292   356    :  12769   Min.   :0.500   Min.   :8.248e+08  
##  67385  :   5088   296    :  12439   1st Qu.:3.000   1st Qu.:9.657e+08  
##  14463  :   3718   480    :  12390   Median :3.500   Median :1.059e+09  
##  68259  :   3228   593    :  12333   Mean   :3.472   Mean   :1.051e+09  
##  27468  :   3218   260    :  12142   3rd Qu.:4.000   3rd Qu.:1.135e+09  
##  19635  :   3016   1196   :  11586   Max.   :5.000   Max.   :1.231e+09  
##  (Other):5516391   (Other):5466292                                      
##     title              genres         
##  Length:5539951     Length:5539951    
##  Class :character   Class :character  
##  Mode  :character   Mode  :character  
##                                       
##                                       
##                                       
## 
\end{verbatim}

@@

\begin{verbatim}
##  [1] -3.0 -2.9 -2.8 -2.7 -2.6 -2.5 -2.4 -2.3 -2.2 -2.1 -2.0 -1.9 -1.8 -1.7 -1.6
## [16] -1.5 -1.4 -1.3 -1.2 -1.1 -1.0 -0.9 -0.8 -0.7 -0.6 -0.5 -0.4 -0.3 -0.2 -0.1
## [31]  0.0  0.1  0.2  0.3  0.4  0.5  0.6  0.7  0.8  0.9  1.0  1.1  1.2  1.3  1.4
## [46]  1.5  1.6  1.7  1.8  1.9  2.0  2.1  2.2  2.3  2.4  2.5  2.6  2.7  2.8  2.9
## [61]  3.0
\end{verbatim}

>>>>

@@

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\subsection{Plot dependency of RMSEs vs Overal Mean Rating
Deviation}\label{plot-dependency-of-rmses-vs-overal-mean-rating-deviation}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{data.frame}\NormalTok{(}\AttributeTok{delta =}\NormalTok{ deviation, }
           \AttributeTok{delta.RMSE =}\NormalTok{ deviation.RMSE) }\SpecialCharTok{|\textgreater{}} 
\FunctionTok{tuning.plot}\NormalTok{(}\AttributeTok{title =} \FunctionTok{TeX}\NormalTok{(r}\StringTok{\textquotesingle{}[RMSE as a function of deviation ($\textbackslash{}delta$) from the Overall Mean Rating ($\textbackslash{}hat\{mu\}$)]\textquotesingle{}}\NormalTok{),}
              \AttributeTok{xname =} \StringTok{"delta"}\NormalTok{, }
              \AttributeTok{yname =} \StringTok{"delta.RMSE"}\NormalTok{, }
              \AttributeTok{xlabel =} \FunctionTok{TeX}\NormalTok{(r}\StringTok{\textquotesingle{}[$\textbackslash{}delta$]\textquotesingle{}}\NormalTok{), }
              \AttributeTok{ylabel =} \StringTok{"RMSE"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-1-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{require}\NormalTok{(edx.capstone.movielens.data)) pak}\SpecialCharTok{::}\FunctionTok{pak}\NormalTok{(}\StringTok{"AzKurban{-}edX{-}DS/edx.capstone.movielens.data"}\NormalTok{)}

\FunctionTok{library}\NormalTok{(edx.capstone.movielens.data)}
\NormalTok{edx }\OtherTok{\textless{}{-}}\NormalTok{ edx.capstone.movielens.data}\SpecialCharTok{::}\NormalTok{edx}
\NormalTok{final\_holdout\_test }\OtherTok{\textless{}{-}}\NormalTok{ edx.capstone.movielens.data}\SpecialCharTok{::}\NormalTok{final\_holdout\_test}

@@

\begin{verbatim}
##      userId         movieId          rating        timestamp            title              genres         
##  Min.   :    1   Min.   :    1   Min.   :0.500   Min.   :7.897e+08   Length:9000055     Length:9000055    
##  1st Qu.:18124   1st Qu.:  648   1st Qu.:3.000   1st Qu.:9.468e+08   Class :character   Class :character  
##  Median :35738   Median : 1834   Median :4.000   Median :1.035e+09   Mode  :character   Mode  :character  
##  Mean   :35870   Mean   : 4122   Mean   :3.512   Mean   :1.033e+09                                        
##  3rd Qu.:53607   3rd Qu.: 3626   3rd Qu.:4.000   3rd Qu.:1.127e+09                                        
##  Max.   :71567   Max.   :65133   Max.   :5.000   Max.   :1.231e+09
\end{verbatim}

@@

\begin{verbatim}
##      userId         movieId          rating        timestamp            title              genres         
##  Min.   :    1   Min.   :    1   Min.   :0.500   Min.   :7.897e+08   Length:999999      Length:999999     
##  1st Qu.:18096   1st Qu.:  648   1st Qu.:3.000   1st Qu.:9.467e+08   Class :character   Class :character  
##  Median :35768   Median : 1827   Median :4.000   Median :1.035e+09   Mode  :character   Mode  :character  
##  Mean   :35870   Mean   : 4108   Mean   :3.512   Mean   :1.033e+09                                        
##  3rd Qu.:53621   3rd Qu.: 3624   3rd Qu.:4.000   3rd Qu.:1.127e+09                                        
##  Max.   :71567   Max.   :65133   Max.   :5.000   Max.   :1.231e+09
\end{verbatim}

@@

\begin{verbatim}
## `summarise()` has grouped output by 'movieId'. You can override using the `.groups` argument.
\end{verbatim}

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-12-1.pdf}

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-16-1.pdf}

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-20-1.pdf}

@@

@@

\begin{verbatim}
## # A tibble: 5 x 5
##   userId `Pulp Fiction (1994)` `Jurassic Park (1993)` `Silence of the Lambs (1991)` `Forrest Gump (1994)`
##    <int>                 <dbl>                  <dbl>                         <dbl>                 <dbl>
## 1     13                     4                     NA                            NA                    NA
## 2     16                    NA                      3                            NA                    NA
## 3     17                    NA                     NA                             5                    NA
## 4     18                     5                      3                             5                    NA
## 5     19                    NA                      1                            NA                     4
\end{verbatim}

@@

\begin{verbatim}
##      userId         movieId          rating        timestamp            title              genres         
##  Min.   :    8   Min.   :    1   Min.   :0.500   Min.   :8.248e+08   Length:1397040     Length:1397040    
##  1st Qu.:18161   1st Qu.: 1022   1st Qu.:3.000   1st Qu.:9.655e+08   Class :character   Class :character  
##  Median :35666   Median : 2109   Median :3.500   Median :1.058e+09   Mode  :character   Mode  :character  
##  Mean   :35902   Mean   : 4616   Mean   :3.473   Mean   :1.051e+09                                        
##  3rd Qu.:53645   3rd Qu.: 3993   3rd Qu.:4.000   3rd Qu.:1.135e+09                                        
##  Max.   :71565   Max.   :65133   Max.   :5.000   Max.   :1.231e+09
\end{verbatim}

@@

\begin{verbatim}
##      userId           movieId            rating        timestamp            title              genres         
##  59269  :   5292   356    :  12769   Min.   :0.500   Min.   :8.248e+08   Length:5539951     Length:5539951    
##  67385  :   5088   296    :  12439   1st Qu.:3.000   1st Qu.:9.657e+08   Class :character   Class :character  
##  14463  :   3718   480    :  12390   Median :3.500   Median :1.059e+09   Mode  :character   Mode  :character  
##  68259  :   3228   593    :  12333   Mean   :3.472   Mean   :1.051e+09                                        
##  27468  :   3218   260    :  12142   3rd Qu.:4.000   3rd Qu.:1.135e+09                                        
##  19635  :   3016   1196   :  11586   Max.   :5.000   Max.   :1.231e+09                                        
##  (Other):5516391   (Other):5466292
\end{verbatim}

@@

\begin{verbatim}
##  [1] -3.0 -2.9 -2.8 -2.7 -2.6 -2.5 -2.4 -2.3 -2.2 -2.1 -2.0 -1.9 -1.8 -1.7 -1.6 -1.5 -1.4 -1.3 -1.2 -1.1 -1.0 -0.9 -0.8 -0.7 -0.6 -0.5 -0.4 -0.3 -0.2 -0.1  0.0  0.1  0.2  0.3  0.4  0.5
## [37]  0.6  0.7  0.8  0.9  1.0  1.1  1.2  1.3  1.4  1.5  1.6  1.7  1.8  1.9  2.0  2.1  2.2  2.3  2.4  2.5  2.6  2.7  2.8  2.9  3.0
\end{verbatim}
